Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-29 examples.tex
simplify tautological implications to save space, but note proof details in appendix. delete obvious invariants that we do not necessarily need to repeat

>>>>

\add{Note that the C presentation in Figure \ref{fig:calltopteinitializeC}
omitted the precondition on the implication of Figure \ref{fig:calltopteinitialize}'s Line \ref{line:page_of_caps},
which is logically equivalent to \textsf{True} since \textsf{entry\_present} checks if the present bit is set in an entry,
and \textsf{pte\_initialize} sets that bit. The actual invariant has this form here, and in the postcondition
of \lstinline|pte_initialize| (Figure \ref{pteinitializespec}), to match the conditional form from earlier in
\lstinline|pte_get_next_table| (which is also provably true when the check of the present bit
determines that the entry was already valid/present).
Our proof discharges the conditional at the join point, rather than eagerly in each branch.
}

### 2025-07-29 examples_C.tex
simplify tautological implications to save space, but note proof details in appendix. delete obvious invariants that we do not necessarily need to repeat

    $\specline{ \ldots \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present entryv})  \urcorner \ast \ulcorner  \textsf{qfrac} = 1 \land \lnot(\textsf{entry\_present (pfn\_set (entryv nextpaddr))}  \urcorner   }_{\rtv}$
    $\specline{ \begin{array}{l}\textsf{entry\_present (pte\_initialized (pfn\_set(entryv nextpaddr)))} \wand  \\ \;\;\;\;\;\;\;\;\  \forall_{i\in \textsf{0 ... 511} } \ldotp  \ghostmaptoken{\textsf{id}}{((\textsf{table\_root (pte\_initialized (pfn\_set(entryv nextpaddr)))}) + \textsf{i * 8})}{\textsf{v-1}} \end{array}  }$ (*\label{line:page_of_capsC}*)

@@

\begin{figure}\footnotesize
  \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
/* @param entry: virtual address of a non-present entry to initialize with a new physical page */
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
void pte_initialize(pte_t *entry) {
  pte_t *local = kalloc(); // Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)
  entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \begin{array}{l} \ulcorner\textsf{entry\_present (pte\_initialize(pfn\_set(entry\_val nextpaddr)),level)}\urcorner \wand \\  \;\;\;\;\;\;\; \forall_{i\in\textsf{0..511}} \; \textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8} \mapsto_{\textsf{id}} \textsf{level-1}    \end{array}  }_{\rtv}$(*\label{line:initialize_post}*)
}
\end{lstlisting}
\vspace{-2em}
\caption{Allocating a physical page }
\label{pteinitializespecC}
\vspace{-1em}
\end{figure}

>>>>

@@

\begin{figure}\footnotesize
  \begin{lstlisting}[language=C,mathescape,escapeinside={(*}{*)}]
/* @param entry: virtual address of a non-present entry to initialize with a new physical page */
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\textsf{vpte,1}} \textsf{entryp entryv} \ast \ghostmaptoken{\textsf{id}}{(\mathsf{entryp})}{\textsf{level}} }$
void pte_initialize(pte_t *entry) {
  pte_t *local = kalloc(); // Allocate a full zeroed page for 512 8-byte entries(*\label{line:call_to_kallocC}*)
  entry->pfn = PTE_ADDR_TO_PFN((uintptr_t) local);(*\label{line:kalloc_install}*)
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \lnot(\textsf{entry\_present entryv}) \ast \textsf{entryp+KERNBASE} \mapsto_{\mathsf{vpte,1}} \mathsf{entryp} \; \textsf{pfn\_set(entryv nextpaddr)}}$
  $\specline{\ghostmaptoken{\delta{}s}{\rtv}{\delta} \ast \forall_{i\in\textsf{0..511}}\ldotp ()\textsf{table\_root (pte\_initialized (pfn\_set (entry\_val nextpaddr))) + i * 8}) \mapsto_{\textsf{id}} \textsf{level-1}   }_{\rtv}$(*\label{line:initialize_post}*)A
}
\end{lstlisting}
\vspace{-2em}
\caption{Allocating a physical page }
\label{pteinitializespecC}
\vspace{-1em}
\end{figure}
